\section{LEADERS}

\hfill

GENERAL RULE:

Leaders aid in melee combat and rally troops. Saxon leaders determine the disposition of Saxon wings. Leaders can be killed, wounded, or captured, affecting victory conditions at the end of the game.

\subsection{Leader Counters}

\subsubsection[Command Radius]{} Each leader has a \textit{command radius} (the leftmost number on the counter). A Saxon leader's command radius is used to determine which combat units are in his wing. If William or Harold fall, the command radius determines which friendly units check morale (see 10.45).

\subsubsection[Rally Range]{} The rightmost number is the leader's \textit{rally range}. The leader can rally any unit in this range as per the rally rules (9.5). A leader with a rally range of "0" can only rally units he is stacked with.

\subsubsection[Subordinate]{} Norman subordinate commanders can only affect units of their own nationality. William can affect Norman units of any nationality.

\subsection{Leaders and Movement}

\subsubsection[Movement]{} Leaders can always move their full movement allowance (6) in any movement segment, regardless of battle order. Leaders ignore the facing rules.

\subsubsection[Enemy ZOCs]{} A leader cannot voluntarily enter an enemy ZOC unless stacked with a friendly combat unit. If the leader is stacked with a friendly unit that is eliminated, leaving the leader in an enemy ZOC, the leader must immediately move three hexes distant from the enemy unit, choosing any path it wishes. However, the leader cannot enter an enemy ZOC or enemy-occupied hex. (The presence of a friendly unit negates enemy ZOC for this purpose). If not retreat path is open, the leader is \textit{captured}.

\subsubsection[Adjacent Enemy Units]{} If an enemy combat unit moves next to a friendly leader that is not stacked with a friendly combat unit, the leader must immediately retreat as in case 10.22.

\subsection{Leaders and Combat}

\subsubsection[Strength]{} If a combat unit is stacked with a leader, it adds \textit{one} to its melee strength (\textit{two} if William or Harold). Leaders do not affect missile fire.

\subsubsection[Disrupted Unit]{} If a disrupted unit is within the rally range of a friendly leader, the owning player subtracts \textit{one} from his rally roll for that unit.

\subsubsection[Routed Unit]{} If a routed unit is within rally range of a friendly leader, the unit is automatically rallied. Thus, William can rally all routed Norman units within two hexes or less from him, while Leofwine can only rally units he is stacked with.

\subsection{Leader Casualties}

\subsubsection[Possible Casualty]{} Any leader in a hex with a combat unit that receives a "1" melee or missile result might become a casualty. The owning player rolls two dice and consults the Leader Casualty Table (10.5), applying any result immediately.

\subsubsection[Killed or Captured]{} Killed or captured leaders are removed from play permanently. (This has little effect, other than the reduced ability to rally and the victory conditions.)

\subsubsection[Wounded]{} A wounded leader remains in the game. The counter is flipped to its back ("ineffective") side. Ineffective leaders move at the rate of two (2) movement points per game turn. They cannot rally units or affect combat.

\subsubsection[Wounded Again]{} Wounded leaders that are wounded again are killed. Shaken leaders that are wounded are wounded, but alive.

\subsubsection[William and Harold]{} If either William or Harold is wounded or killed, all friendly units within command radius must make an immediate morale check.

\subsection{The Leader Casualty Table}

\begin{tabular}{ |ccl| }
  \hline & &\\ [-2.0ex]
  \textbf{Fire} & \textbf{Melee} & \textbf{Result} \\
  \hline & &\\ [-2.0ex]
  2,12 & 2,12 & Leader Killed \\
  3,11 & 3 & Leader Wounded\textsuperscript{1} \\
  -- & 11,4,10 & Leader Wounded\textsuperscript{2} \\
  -- & 5,9 & Leader Shaken\textsuperscript{3} \\
  4-10 & 6,7,8 & Nothing Happens \\
  \hline
\end{tabular}

\textsuperscript{1:} Leader remains in current status. Both his command radius and rally radius are reduced by 1 for the rest of the game.

\textsuperscript{2:} Leader is flipped to his ineffective side for the remainder of the Assault Period.

\textsuperscript{3:} Leader is unhorsed/shaken. Flip the leader to his ineffective side for the remainder of this Battle Phase and all of the next. If already shaken, roll 1d6: 1-3 killed, 4-6 nothing happens.