\section{VICTORY CONDITIONS}

\subsection{How to Win}

The strategic objective of the Normans was to reach and take London. Harold's army blocked the only road leading from William's landing site to the capital. If the Normans had been stopped, Harold would have kept his kingdom.

\subsubsection[Norman Victory]{} The Norman player wins a Strategic Victory if, at any time, he controls all the road hexes on Senlac Hill (0811 to 0116). Control is defined as being the last to occupy or pass through a hex.

If the Normans have lost more casualty points than the Saxons, then the Normans win only a Tactical Victory (their army is probably not strong enough to continue the campaign).

\subsubsection[Norman Victory Alt]{} The Norman player also wins a Strategic Victory if he eliminates all the Saxon housecarle units.

However, if the Normans have lost more casualty points than the Saxons, the game is a Draw (the armies accomplish nothing but wear each other out).

\subsubsection[Saxon Victory]{} The Saxons win a Strategic Victory if the Normans fail to accomplish either of their victory conditions by the end of the Second Assault Period (they have stopped William's invasion in its tracks).

\subsection{Casualty Points}

Each time a unit is eliminated, its side accumulates casualty points. These are totalled at the end of the game, and can affect possible Norman victories (12.1). Points are awarded on the given table.

\begin{center}
  \begin{tabular}{ l c|c }
    \textbf{Unit} & \textbf{Casualty Points} \\
    \hline
    "A" Morale unit & 15 \\
    "B" Morale unit & 7 \\
    "C" Morale unit & 3 \\
    "D" Morale unit & 2 \\
    "E" Morale unit & 1 \\
    William & 50 \\
    Harold & 25 \\
    Odo & 10 \\
    Other Leader & 5
  \end{tabular}
\end{center}